---
title: "FOXP1 Analysis"
author: "Saffron"
date: "2023-07-27"
categories: [analysis, code, visualizations]
---

Just messing around with interactive plots.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height=5, fig.width=7, message = F, warning = F)
```

```{r packages}
library(pacman)
p_load(readr, tidyr, magrittr, knitr, tidyverse, janitor, broom, ggplot2, forcats, epitools, dplyr, xfun, plotly)
```

The data used for this tinkering is found on the National Library of Medicine [ClinVar database](https://www.ncbi.nlm.nih.gov/clinvar/?term=FOXP1%5Bgene%5D&redir=gene).   

```{r}
# load data
# imported from directory
# Helpful note to self: when loading csv files from "copy path" delete 'C:' and change all \ to /
CVFP1 <- read_csv("/Users/mckjo/OneDrive/Desktop/GitBlog/posts/FOXP1 Analysis/ShinyClinVar.csv")

# clean variable names
janitor::clean_names(CVFP1)

# number of rows (144)
nrow(CVFP1)

# convert character variables to factors
CVFP1 <- CVFP1 %>%
  mutate_if(is.character, as.factor)

# summary of dataset 
summary(CVFP1)

# removing the first two colummns   
CVFP1 <- CVFP1[, -c(1, 2)] 

# double checking the two columns were removed
head(CVFP1)

# list of column names 
(column_names <- colnames(CVFP1))
```

## Static Data Visualizations

### Bar Plots

```{r}
ggplot(CVFP1, aes(x = ClinicalSignificance)) + geom_bar(color = "black", fill = "lightgray") + 
  labs(title = "Clinical Significance of Genetic Variant from FOXP1", 
       subtitle = "n = 144",
       x = "Level of Significance", 
       y = "Number of Individuals") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
```

## Interactive Data Visualizations

### Bar Plots

**Note**: As of August 3rd, 2023, I am unable to render using 'plotly'. The code chunk below has came to a halt -- I've troubleshooted by restarting, clearing the console and environment, checking for package conflicts, making sure everything is updated, removed my previous code with `layout()` and am still having this error:

* checked again on August 21st   


Quitting from lines 68-138 \[plotly\] (index.qmd)\
Error in `add_html_caption()`:\
! unused argument (xfun::grep_sub("[^1]\* \<\[\^\>\]+aria-labelledby\[ \]\* =\[ \] \*  "(\[\^ "\]+) ".\*\$"," \\1", x))\
Backtrace:\
1. global .main()\
2. execute(...)\
3. rmarkdown::render(...)\
4. knitr::knit(knit_input, knit_output, envir = envir, quiet = quiet)\
5. knitr:::process_file(text, output)\
...\
14. sew(res, options)\
15. knitr:::sew.list(x, options, ...)\
16. base::lapply(x, sew, options, ...)\
17. FUN(X\[\[i\]\], ...)\
18. knitr:::sew.knit_asis(x, options, ...)\
Execution halted

[^1]: \^\<

```{r plotly}
# Clinical Significance Bar Plot

(df.bplot <- CVFP1 %>% dplyr::group_by(ClinicalSignificance) %>% dplyr::count())

y.cs <- c(8, 5, 31, 20, 33, 2, 45)

text.clin.sig <- c("8", "5", "31", "20", "33", "2", "45")

## showing the counts of cases based on Clinical Significance 
# b.plot <- plotly::plot_ly(data = df.bplot, 
#                          x = ~ClinicalSignificance, 
#                          y = ~y.cs, 
#                          type = "bar", 
#                          text = text.clin.sig, 
#                          marker = list(color = "maroon")) 

# b.plot

## showing the percentage of case based on Clinical Significance 

### Benign  
ben <- paste(round(8/144*100, digits = 2), "%", sep = "")

### Benign/Likely Benign  
ben.lben <- paste(round(5/144*100, digits = 2), "%", sep = "")

### Likely Benign  
lben <- paste(round(31/144*100, digits = 2), "%", sep = "")

### Likely Pathogenic  
lpat <- paste(round(20/144*100, digits = 2), "%", sep = "")

### Pathogenic   
pat <- paste(round(33/144*100, digits = 2), "%", sep = "")

### Pathogenic/Likely Pathogenic  
pat.lpat <- paste(round(2/144*100, digits = 2), "%", sep = "")

### Uncertain/Conflicting 
unc.con <- paste(round(45/144*100, digits = 2), "%", sep = "")


text.per <- c(ben, ben.lben, lben, lpat, pat, pat.lpat, unc.con)

# barplot with percentages 
# b.per.plot <- plotly::plot_ly(data = df.bplot, 
#                          x = ~ClinicalSignificance, 
#                          y = ~y.cs, 
#                          type = "bar", 
#                          text = text.per, 
#                          marker = list(color = "skyblue")) 

# b.per.plot

# Consequence Bar Plot 

(df.c.bplot <- CVFP1 %>% dplyr::group_by(consequence) %>% dplyr::count())

# showing the counts 
y.con <- c(1, 21, 71, 13, 17, 21)

# c.b.plot <- plotly::plot_ly(data = df.c.bplot,
#                          x = ~consequence,
#                          y = ~y.con, 
#                          type = "bar", 
#                          marker = list(color = "lavender", 
#                                        line = list(color = "darkgrey", 
#                                                    width = 1.5))) 

# c.b.plot
```

## Interactive Plots Using 'highcharter'    

**Note**: As of August 5th, `plotly` is still not running correctly so I'm going to try out this package `highcharter`    

My ideas and inspiration came from this (site)[https://www.datanovia.com/en/lessons/highchart-interactive-bar-plot-in-r/]    

* August 21st: same code warning with "highcharter" as it is with "plotly" - I figure at this point it's good to show my flops as well as my successes, so enjoy this little flop   

```{r highcharter}
# install.packages("highcharter")   
library(highcharter) 

# setting highcharter options 
options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))
```

```{r basic barplots}
# basic <- df.c.bplot %>% 
#  highcharter::hchart("column", 
#         hcaes(
#           x = consequence, 
#           y = y.con
#         ), 
#         color = "lavender", 
#         borderColor = "black", 
#         pointWidth = 90)
# basic
```

## Interactive Plots Using `gganimate`    

My ideas and inspiration came from this (site)[https://towardsdatascience.com/create-animated-bar-charts-using-r-31d09e5841da] 


```{r gganimate}
# install.packages("gganimate")  
library(gganimate)
```


```{r}
cvfp1_formatted <- CVFP1 %>%
  group_by(consequence) %>%
  # The * 1 makes it possible to have non-integer ranks while sliding
  mutate(C.S.Pathogenic = (ClinicalSignificance=="Pathogenic")) 
```

```{r staticplot}
staticplot <- ggplot(CVFP1, aes(x = ClinicalSignificance)) + geom_bar(color = "black", fill = "lightgray") + 
  labs(title = "Clinical Significance of Genetic Variant from FOXP1", 
       subtitle = "n = 144",
       x = "Level of Significance", 
       y = "Number of Individuals") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
```

### Coding a Plot.gif   

```{r animation}
aniplot <- staticplot + 
  transition_states(ClinicalSignificance)

### Note to self: 

# do not print/view 'aniplot' at this stage
    # every static plot gets its own .png file (101 pngs here)

# install.packages("gifski")
library(gifski)

animate(aniplot, 200, fps = 20, width = 1200, height = 1000, 
        renderer = gifski_renderer("gganim.gif"))
```













